---
title: "p8105_hw2_js5431"
author: "J Shearston"
date: "September 25, 2018"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem #1

#### Importing and tidying up the data

```{r import and look at data, message = FALSE}
nyc_transit = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names()

#View(nyc_transit)
#skimr::skim(nyc_transit)
```

#### Describing the data

*Instructions:* Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

*Response:* 

#### Question responses

*Item: How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St A/B/C/D; 125st 1; 125st 4/5); the distinct function may be useful here.*

*Item: How many stations are ADA compliant?*

*Item: What proportion of station entrances / exits without vending allow entrance?*

*Item: Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? How many are ADA compliant?*
