---
title: "p8105_hw2_js5431"
author: "J Shearston"
date: "September 25, 2018"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))

```

## Problem #1

#### Importing and tidying up the data

```{r import and clean transit data, message = FALSE}
nyc_transit = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:station_longitude, entry, entrance_type, vending, ada, starts_with("route")) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE)) %>% 
  rename(station_lat = station_latitude, 
         station_long = station_longitude)
#nyc_transit

```

#### Describing the data

*Instructions:* Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

*Response:* This dataset contains information on New York City subway lines and stations, including the line (e.g. 4th Avenue), station name, exact station location (latitute and longitude), routes served (e.g. A, 1), entrances and entrance types, presence of vendors, and ADA compliance. The data have been imported to R, variable names have been cleaned, variables have been re-ordered to make them easier to read, and the "entry" variable has been changed from a character to a logical variable. In its current form, the data is in a table of dimensions `r nrow(nyc_transit)` x `r ncol(nyc_transit)`. The data are somewhat tidy, but more work could be done. I would suggest creating an additional variable that sums the number of routes served by a station.  

#### Question responses

**Item: How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St A/B/C/D; 125st 1; 125st 4/5); the distinct function may be useful here.**

There are `r nrow(distinct(nyc_transit, station_name, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11))` distinct stations in this dataset. 

**Item: How many stations are ADA compliant?**

There are `r tally(nyc_transit, ada == TRUE)` ADA compliant stations.

**Item: What proportion of station entrances / exits without vending allow entrance?**

`r filter(nyc_transit, entry == TRUE & vending == "NO") %>% tally(entry)` out of `r tally(nyc_transit, vending == "NO")` station entrances/exits without venidng allow entrance.

**Item: Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? How many are ADA compliant?**

Reformatting

```{r}

```

`r ` distinct stations serve the A train. Of those, `r` are ADA compliant.


## Problem #2

#### Reading in and cleaning Mr. Trash Wheel data

```{r import and clean trash wheel data}
trash_wheel = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                                 sheet = "Mr. Trash Wheel", 
                                 range = "A2:N338") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(dumpster)) %>% 
  mutate(
    sports_balls = round(sports_balls, digits = 0),
    sports_balls = as.integer(sports_balls)
    )
```

#### Reading in and cleaning precipitation data

2017 Precipitation Data

```{r clean precip 2017 data}
precip_2017 = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                                 sheet = "2017 Precipitation", 
                                 range = "A2:B14") %>% 
  janitor::clean_names() %>%
  rename(precip_inches = total) %>% 
  mutate(year = "2017",
         month = month.name) %>% 
  filter(!is.na(precip_inches))

View(precip_2017)
```

 2016 Precipitation Data
 
```{r clean precip 2016 data}
precip_2016 = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                                 sheet = "2016 Precipitation", 
                                 range = "A2:B14") %>% 
  janitor::clean_names() %>%
  rename(precip_inches = total) %>% 
  mutate(year = "2016",
         month = month.name)

View(precip_2016)

```

Combining 2016 and 2017 Precipitation Data

```{r binding precip data}
precip = bind_rows(precip_2016, precip_2017)
View(precip)
summary(precip)
```

#### Description of the Mr. Trash Wheel dataset

*Instructions: Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2017? What was the median number of sports balls in a dumpster in 2016?*

The Mr. Trash Wheel dataset provides a summary of the amount of trash collected by the Mr. Trash Wheel harbor cleaning device, and has `r nrow(trash_wheel)` rows and `r ncol(trash_wheel)` columns. Information is organized by dumpster number, and Key variables include date specific information such as year, month, and actual date, volume and weight of trash, and a series of variables about specific types of trash (e.g. plastic bottles, cigarette butts, sports balls). Finally, the number of homes powered by the collected trash is also included (woohoo!).

The precipiation dataset includes precipitation in inches from 2016 and 2017. There were `r nrow(precip)` rows and `r ncol(precip)` columns in the dataset, including information from January of 2016 to December of 2017. Key variables included year (2016 or 2017), precipiation in inches, and month.

The total precipitation in 2017 for January-December was `r filter(precip, year == "2017") %>% summarise(sum_precip_2017 = sum(precip_inches))` inches. In 2016, the median number of sports balls accumulated by Mr. Trash Wheel was `r filter(trash_wheel, year == "2016") %>% summarise(med_balls_2016 = median(sports_balls))`.

## Problem #3

#### Loading and cleaning BRFSS data

```{r loading & cleaning brfss data}
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)

brfss = p8105.datasets::brfss_smart2010 %>% 
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>% 
  select(year, locationabbr, locationdesc, response, data_value) %>% 
  spread(key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  mutate(ex_vg = excellent + very_good) %>% 
  select(year, locationabbr, locationdesc, ex_vg, excellent,
         very_good, good, fair, poor) # re-ordering for my benefit
  
```

#### Answering / doing four items

*Item 1: How many unique locations are included in the dataset? Is every state represented? What state is observed the most?*

* `r summarise(brfss, unique_loc = n_distinct(locationdesc))` unique locations are included in the dataset.

* Every state is represented.
```{r}
summarise(brfss, unique_state = n_distinct(locationabbr))
```

* Florida is the most observed state.
```{r}
table(brfss$locationabbr)
```


*Item 2: In 2002, what is the median of the “Excellent” response value?*

The median "Excellent" response in 2002 is `r filter(brfss, year == "2002") %>% summarise(ex_median = median(excellent, na.rm = TRUE))`.


*Item 3: Make a histogram of “Excellent” response values in the year 2002.*

```{r}
filter(brfss, year == "2002") %>% 
  ggplot(aes(x = excellent)) + 
  geom_histogram() +
  labs(
    title = "'Excellent' Self-Reported Overall Health, BRFSS 2002",
    x = "Proportion of respondents who reported 'excellent' overall health",
    y = "Number of counties",
    caption = "Data from the BRFSS, p8105 datasets package") + 
  scale_x_continuous(breaks = c(5, 10, 15, 20, 25, 30, 35, 40, 45),
                   labels = c(5, 10, 15, 20, 25, 30, 35, 40, 45)) + 
  scale_y_continuous(breaks = c(0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20),
                   labels = c(0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20))
```


*Item 4: Make a scatterplot showing the proportion of “Excellent” response values in New York County and Queens County (both in NY State) in each year from 2002 to 2010.*

```{r}
filter(brfss, 
       locationdesc == "NY - New York County" | locationdesc == "NY - Queens County") %>% 
  ggplot(aes(x = year, y = excellent, color = locationdesc)) +
  geom_point(alpha = .75, size = 2) +
  geom_smooth(se = FALSE) +
  labs(
    title = "'Excellent' Self-Reported Health in 2 NY Counties",
    x = "Year",
    y = "Proportion of respondents",
    caption = "Data from the BRFSS, p8105 datasets package") + 
  scale_color_hue(name = "County")

```


